```{r, libs, message=FALSE}
library(readxl)
library("kableExtra")
library(tidyverse)
library(gridExtra)
library(reshape2)
library(dplyr)
library(edgeR)
library(clusterProfiler)
library(BiocGenerics)
library(GenomicFeatures)
library(Biostrings)
library(rtracklayer)
library(data.table)
```

```{r}
pc.gtf <- rtracklayer::import("https://ftp.ncbi.nlm.nih.gov/genomes/genbank/protozoa/Phytophthora_cinnamomi/latest_assembly_versions/GCA_018691715.1_ASM1869171v1/GCA_018691715.1_ASM1869171v1_genomic.gtf.gz")
pc.df <- as.data.frame((pc.gtf))

pc.genes <- filter(pc.df, type == "start_codon")
```




download .faa file https://ftp.ncbi.nlm.nih.gov/genomes/genbank/protozoa/Phytophthora_cinnamomi/latest_assembly_versions/GCA_018691715.1_ASM1869171v1/GCA_018691715.1_ASM1869171v1_protein.faa.gz

```{r}
#setwd("~/projects/Schroeter_Pcinnamomi/PcEffector")

# Read the .faa file
faa_file <- "GCA_018691715.1_ASM1869171v1_protein.faa"  
sequences <- readAAStringSet(faa_file, format = "fasta")

formatted_sequences <- paste0(">", names(sequences), "\n", as.character(sequences), "\n")
output_file <- "/mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector/Pcgenes.fasta" 
writeLines(formatted_sequences, output_file)

Pcgenenames <- read.table('pcgeneids.tsv', row.names = 1, quote = "", sep='\t', fill = TRUE, header = TRUE)

load("Pceffetor.RData")

# first rearrange cols
allgenes <- detectedgenes
allgenes$gene_id<-rownames(allgenes)
rownames(allgenes)<-seq(nrow(allgenes))
head(allgenes)

Pcgenenames <- read.table('pcgeneids.tsv', row.names = 1, quote = "", sep='\t', fill = TRUE, header = TRUE)

pcprotein.id <- dplyr::right_join(Pcgenenames, allgenes, by = 'gene_id')

# Define a list of protein IDs to extract
protein_list <- pcprotein.id$protein_id

```

```{r}
# Read the .faa file
faa_file <- "GCA_018691715.1_ASM1869171v1_protein.faa"  
sequences <- readAAStringSet(faa_file)

# Extract headers and sequences
headers <- names(sequences)
sequence_strings <- as.character(sequences)

# Extract protein identifiers from headers
protein_identifiers <- sub("^>(\\S+).*", "\\1", headers)

# Create a data frame with protein identifiers and sequences
df <- data.frame(protein_id = protein_identifiers, Sequence = sequence_strings, stringsAsFactors = FALSE)

# Filter the "ProteinID" column to include only the first 12 characters
df$protein_id <- substr(df$protein_id, 1, 12)

# Print the updated data frame
head(df)
uniqueprot <- unique(df[df$protein_id,])
pcprotein.idfaa <- dplyr::right_join(df, pcprotein.id, by = 'protein_id')
pcprotein.idfaa <- pcprotein.idfaa[, 1:4]
head(pcprotein.idfaa)
```
function used to create .fasta for inputs

```{r}
#function used to make .fasta files for analysis
write_fasta <- function(df, file_path) {
  fasta_string <- paste0(">", df$protein_id, "\n", df$Sequence, "\n")
  writeLines(fasta_string, con = file_path)
}

generate_fasta_file <- function(df, output_file) {
  write_fasta(df, output_file)
}
```

```{r}
testfasta <- head(pcprotein.idfaa)
output_file <- "/mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector/Raw_fasta/test.fasta"

generate_fasta_file(testfasta, output_file)
#use for testing terminal scripts (InterProScan, SignalP and EffectorP) - 6 genes
```

```{r}
#Pc genes detect in transcriptome -- 13595 genes detected 
output_file <- "/mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector/Raw_fasta/detected_genes.fasta"

generate_fasta_file(pcprotein.idfaa, output_file)

```

```{r}
#all Pc genes from .faa file -- 21335 genes 
output_file <- "/mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector/Raw_fasta/allpcgenes.fasta"

generate_fasta_file(df, output_file)

```

#######################
 InterProScan analysis
#######################

barry@ziemann-01:/mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector/interproscan/interproscan-5.62-94.0$ ./interproscan.sh -appl CDD,PANTHER,Pfam,PROSITEPATTERNS,PROSITEPROFILES,SFLD,SMART,TIGRFAM,Phobius,TMHMM,SUPERFAMILY -i /mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector/Raw_fasta/allpcgenes.fasta -iprlookup --goterms -b /mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector/interproscan_output/
05/06/2023 12:56:46:576 Welcome to InterProScan-5.62-94.0
05/06/2023 12:56:46:577 Running InterProScan v5 in STANDALONE mode... on Linux
05/06/2023 12:56:51:785 RunID: ziemann-01_20230605_125651556_b2fz
05/06/2023 12:57:02:847 Loading file /mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector/Raw_fasta/allpcgenes.fasta
05/06/2023 12:57:02:848 Running the following analyses:
[CDD-3.20,NCBIfam-11.0,PANTHER-17.0,Pfam-35.0,Phobius-1.01,ProSitePatterns-2022_05,ProSiteProfiles-2022_05,SFLD-4,SMART-9.0,SUPERFAMILY-1.75,TMHMM-2.0c]
Available matches will be retrieved from the pre-calculated match lookup service.

Matches for any sequences that are not represented in the lookup service will be calculated locally.
05/06/2023 12:57:07:851 Uploaded 20850 unique sequences for analysis
05/06/2023 12:59:50:492 87% completed
05/06/2023 13:59:50:678 19% completed
 05/06/2023 14:59:50:824 24% completed
05/06/2023 15:59:51:007 77% completed
05/06/2023 16:12:19:464 100% done:  InterProScan analyses completed 

```{r}
interproscanresults <- read_tsv("allpcgenes.fasta.tsv")

colnames(interproscanresults) <- c("protein_id", "protein_md5", "protein_length", 
                     "analysis", "signature_id", 
                     "signature_desc", "start", "end", 
                     "score", "status", "date", 
                     "interpro_id", "interpro_desc", "go_terms")
head(interproscanresults)

uniqueips <- distinct(interproscanresults, protein_id)

```

```{r}
#extract signal peptides from interproscan results 
interproscanSP <- unique(interproscanresults[interproscanresults$signature_id == "SIGNAL_PEPTIDE",])

#Phobius predicts 3050 proteins with predicted signal peptides from 16957 proteins scanned - less than 80 amino acids not scanned by interProScan
```

```{r}
interproscanTMHMM <- unique(interproscanresults[interproscanresults$signature_id == "TMhelix",])

filtered_interproscanTMHMM <- interproscanTMHMM %>%
  group_by(protein_id) %>%
  filter(n() <= 1) %>% 
  filter(end <= 60)

#TMHMM predicts 594 proteins that aren't considered transmembrane

filtered_interproscanTMHMM.id <- dplyr::right_join(Pcgenenames, filtered_interproscanTMHMM, by = 'protein_id')
```

Tmhmm v.2.0 was used to predict transmembrane proteins from the combined predicted secretome (Krogh et al., 2001). If a protein contains at least one transmembrane domain after the first 60 amino acids or at least two transmembrane domains in total, it was labelled as a putative transmembrane protein.

Ranking effector candidates according to their expression during infection with RNA-sequencing (RNA-seq) data or through structural similarity to known effectors are recommended practices for effector candidate prioritization.

Sperschneider, J., Gardiner, D.M., Dodds, P.N., Tini, F., Covarelli, L., Singh, K.B., Manners, J.M. and Taylor, J.M. (2016), EffectorP: predicting fungal effector proteins from secretomes using machine learning. New Phytol, 210: 743-761. https://doi.org/10.1111/nph.13794


##################
 SignalP analysis
##################

setwd("~/projects/Schroeter_Pcinnamomi/PcEffector/signalp-6.0/signalp6_fast/")
Sys.setenv(PATH = paste("/home/barry/.local/bin", Sys.getenv("PATH"), sep = ":"))

SIGNALP_DIR <- system("python3 -c 'import signalp; import os; print(os.path.dirname(signalp.__file__))'", intern = TRUE)

Sys.getenv("PATH")

system("signalp6 -h")

system("signalp6 --fastafile ~/projects/Schroeter_Pcinnamomi/PcEffector/Raw_fasta/allpcgenes.fasta --organism eukarya --output_dir ~/projects/Schroeter_Pcinnamomi/PcEffector/SignalPall_output --format none --mode fast")

system("signalp6 --fastafile ~/projects/Schroeter_Pcinnamomi/PcEffector/Raw_fasta/allpcgenes.fasta --organism eukarya --output_dir ~/projects/Schroeter_Pcinnamomi/PcEffector/SignalPall_output --format none --mode fast")
Predicting: 100%|██████████| 21334/21334 [39:18<00:00,  9.05sequences/s]

# run on signalP_6.0 and import results 

```{r}
# Read the prediction_results.txt file into a data frame
prediction_results <- read.table("prediction_results.txt", header = F, comment.char = "#", sep = "\t")
colnames(prediction_results) <- c("protein_id", "Prediction", "OTHER", "SP(Sec/SPI)", "CS Position")
head(prediction_results)

# Filter the prediction_results data frame by "SP" in the Prediction column
signalPsecreted <- filter(prediction_results, Prediction == "SP")

# Use separate to split CS Position column
split_results <- signalPsecreted %>%
  separate(`CS Position`, into = c("cleavage_site", "probability"),
           sep = "\\. ", remove = FALSE)

signalPSP <- split_results %>%
  mutate(cleavage_site = gsub("CS pos: ", "", cleavage_site),
         probability = gsub("Pr: ", "", probability))

head(signalPSP)

#1878/21334 proteins predicted to contain a Sec/SPI SPs using signalP6.0 

split_results$probability <- as.numeric(signalPSP$probability)
hist(split_results$probability, main = "Probability Histogram", xlab = "Probability")

#filter out low probability SP
signalPSP.filtered <- subset(signalPSP, probability >= 0.7)

#1694/1878 with a probability of 0.7 or greater

```

```{r}
#list of proteins from interproscan - phobius/tmhmm and signalP predicted to be secreted
phobius.tmhmm.signalp <- unique(c(interproscanSP$protein_id, filtered_interproscanTMHMM$protein_id, signalPSP$protein_id))
pts <- df[df$protein_id %in% phobius.tmhmm.signalp, ]

pts.id <- dplyr::right_join(Pcgenenames, pts, by = 'protein_id')

```

```{r}
#minus those that are OTHER in signalP (removes some ribosomal proteins)
signalPnon <- filter(prediction_results, Prediction == "OTHER")
# Find the common protein IDs
common_protein_ids <- intersect(interproscanSP$protein_id, signalPnon$protein_id)

# Remove the common protein IDs from the dataset
interproscanSP <- interproscanSP[!interproscanSP$protein_id %in% common_protein_ids, ]

```

```{r}
#join interproscan and signalP results together use this file for effectorP analysis
interproscan_protein_ids <- interproscanSP$protein_id
signalP_protein_ids <- signalPSP$protein_id

combined_protein_ids <- c(interproscan_protein_ids, signalP_protein_ids)

totalSP <- (unique(combined_protein_ids))

SPdf <- df[df$protein_id %in% totalSP, ]

#1878 genes predicted to have signal peptides
```

###############################
 Effectors from the literature 
###############################

```{r}
#PFAM accessions associated with pathogenicity

#elicitin Pfam domain PF00964
PF00964 <- unique(interproscanresults[interproscanresults$signature_id == "PF00964",])
#serine protease inhibitor Kazal-like domain PF07648
PF07648 <- unique(interproscanresults[interproscanresults$signature_id == "PF07648",])
#chymotrypsin PF00089 
PF00089 <- unique(interproscanresults[interproscanresults$signature_id == "PF00089",])
#PAN/Apple domain PF14295 
PF14295 <- unique(interproscanresults[interproscanresults$signature_id == "PF14295",])
#Cathepsin propeptide inhibitor PF08246
PF08246 <- unique(interproscanresults[interproscanresults$signature_id == "PF08246",])
# Lytic polysaccharide mono-oxygenase PF03067
PF03067 <- unique(interproscanresults[interproscanresults$signature_id == "PF03067",])
#Transglutaminase_elicitor PF16683 
PF16683 <- unique(interproscanresults[interproscanresults$signature_id == "PF16683",])
#Cysteine_rich_secretory_protein_family PF00188 
PF00188 <- unique(interproscanresults[interproscanresults$signature_id == "PF00188",])
#pectin degradation (PF03283, PF00544, and PF03211)
pectin_degradation <- unique(interproscanresults[interproscanresults$signature_id == unique(c("PF03283","PF00544","PF03211")),])

pfameffectors <- as.data.frame(c(PF00964$protein_id, PF07648$protein_id, PF00089$protein_id, PF14295$protein_id, PF08246$protein_id, PF03067$protein_id, PF16683$protein_id, PF00188$protein_id, pectin_degradation$protein_id))

names(pfameffectors)[1] <- "protein_id"

pfameffectors.id <- dplyr::right_join(Pcgenenames, pfameffectors, by = 'protein_id')

```

Get supplementary files from https://rdcu.be/dcczO
https://static-content.springer.com/esm/art%3A10.1186%2Fs12864-021-07552-y/MediaObjects/12864_2021_7552_MOESM2_ESM.xls

```{r, Engelbrecht supp rxlr}
#only considered the 181 candidate RxLR -IUM83_06578-RA/RB alternatively spliced total 180 rxlr

pcrxlr <- read_xls('12864_2021_7552_MOESM2_ESM.xls', 
                   sheet = 'TableS2',
                   col_names = F,
                   range = 'A3:A183') 
# Rename the column
colnames(pcrxlr) <- c("gene_id")

# Remove the last three characters of each row
pcrxlr$gene_id <- substr(pcrxlr$gene_id, 1, nchar(pcrxlr$gene_id) - 3)

as.data.frame(pcrxlr)

rxlr.df <- Pcgenenames %>%
  filter(gene_id %in% pcrxlr$gene_id)

head(rxlr.df)

```

```{r, Engelbrecht supp crn}
#49 CRNs (45 genes, of which four have alternative transcript variants)

pccrn <- read_xls('12864_2021_7552_MOESM2_ESM.xls', 
                   sheet = 'TableS3',
                   col_names = F,
                   range = 'A3:A51') 
# Rename the column
colnames(pccrn) <- c("gene_id")

# Remove the last three characters of each row
pccrn$gene_id <- substr(pccrn$gene_id, 1, nchar(pccrn$gene_id) - 3)

as.data.frame(pccrn)

crn.df <- Pcgenenames %>%
  filter(gene_id %in% pccrn$gene_id)

head(crn.df)
```

```{r, Engelbrecht supp nlp}
#61 necrosis-inducing proteins (NLPs)

pcnlp <- read_xls('12864_2021_7552_MOESM2_ESM.xls', 
                   sheet = 'TableS4',
                   col_names = F,
                   range = 'A3:A63') 
# Rename the column
colnames(pcnlp) <- c("gene_id")

# Remove the last three characters of each row
pcnlp$gene_id <- substr(pcnlp$gene_id, 1, nchar(pcnlp$gene_id) - 3)

as.data.frame(pcnlp)

nlp.df <- Pcgenenames %>%
  filter(gene_id %in% pcnlp$gene_id)

head(nlp.df)
```

#####################
 EffectorP analysis 
#####################

```{r}
#list of know effectors based on published data 
knowneffectors <- as.data.frame(c(nlp.df$protein_id, crn.df$protein_id, rxlr.df$protein_id))

names(knowneffectors)[1] <- "protein_id"
#add known effectors to predicted effectors

allsecreted <- unique(c(nlp.df$protein_id, crn.df$protein_id, rxlr.df$protein_id, SPdf$protein_id))
allsecreted <- as.data.frame(allsecreted)
names(allsecreted) <- "protein_id"

allsecreted.id <- dplyr::right_join(df, allsecreted, by = 'protein_id')

output_file <- "/mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector//EffectorP_3.0.0-beta/allsecreted.fasta"

generate_fasta_file(allsecreted.id, output_file)

detectedsecreted.id <- dplyr::right_join(pcprotein.idfaa, allsecreted, by = 'protein_id')

#path to where the 
setwd("~/projects/Schroeter_Pcinnamomi/PcEffector/EffectorP_3.0.0-beta/")

system("/home/barry/.virtualenvs/my-python/bin/python EffectorP.py -i allsecreted.fasta > allsecreted_effectorPresults.txt")

```

1878 proteins were provided as input in the following file: allSP.fasta

Number of predicted effectors: 741
Number of predicted cytoplasmic effectors: 379
Number of predicted apoplastic effectors: 362

39.5 percent are predicted effectors.
20.2 percent are predicted cytoplasmic effectors.
19.3 percent are predicted apoplastic effectors.

6.3 percent of the cytoplasmic effectors are predicted dual-localized (cytoplasmic/apoplastic).
11.3 percent of the apoplastic effectors are predicted dual-localized (apoplastic/cytoplasmic).

```{r}
effectorp.results <- read.delim("allSP_effectorPresults.txt", sep = "\t", stringsAsFactors = FALSE, skip = 10)
effectorp.results <- effectorp.results[1:(nrow(effectorp.results) - 15), ]
colnames(effectorp.results) <- c("protein_id", "Cytoplasmic_effector", "Apoplastic_effector", "Non_effector", "Prediction")
effectorp.results$protein_id <- gsub(" ", "", effectorp.results$protein_id)

head(effectorp.results)

# make this all non effectors from effectorP
predictedeffectors <- subset(effectorp.results, Prediction != "Non-effector")
predictedeffectors.id <- dplyr::right_join(pcprotein.idfaa, predictedeffectors, by = 'protein_id')
```

```{r}
output_file <- "/mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector//EffectorP_3.0.0-beta/phobtmhmmsigp.fasta"

generate_fasta_file(pts, output_file)

#path to where the 
setwd("~/projects/Schroeter_Pcinnamomi/PcEffector/EffectorP_3.0.0-beta/")

system("/home/barry/.virtualenvs/my-python/bin/python EffectorP.py -i phobtmhmmsigp.fasta > phobtmhmmsigpeffectorp.txt")
```

3383 proteins were provided as input in the following file: phobtmhmmsigp.fasta

Number of predicted effectors: 1241
Number of predicted cytoplasmic effectors: 845
Number of predicted apoplastic effectors: 396

36.7 percent are predicted effectors.
25.0 percent are predicted cytoplasmic effectors.
11.7 percent are predicted apoplastic effectors.

3.9 percent of the cytoplasmic effectors are predicted dual-localized (cytoplasmic/apoplastic).
10.6 percent of the apoplastic effectors are predicted dual-localized (apoplastic/cytoplasmic).

```{r}
ptseffectorp.results <- read.delim("EffectorP_3.0.0-beta/phobtmhmmsigpeffectorp.txt", sep = "\t", stringsAsFactors = FALSE, skip = 10)
ptseffectorp.results <- ptseffectorp.results[1:(nrow(ptseffectorp.results) - 15), ]
colnames(ptseffectorp.results) <- c("protein_id", "Cytoplasmic_effector", "Apoplastic_effector", "Non_effector", "Prediction")
ptseffectorp.results$protein_id <- gsub(" ", "", ptseffectorp.results$protein_id)

head(ptseffectorp.results)

# make this all non effectors from effectorP
predictedeffectors2 <- subset(ptseffectorp.results, Prediction != "Non-effector")
predictedeffectors2.id <- dplyr::right_join(pts.id, predictedeffectors2, by = 'protein_id')

#make .fasta for PHI database
output_file <- "/mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector/Raw_fasta/predictedeffectors.fasta"

generate_fasta_file(predictedeffectors2.id, output_file)
```

###
PHI database analysis
###

run predictedeffectors.fasta at http://phi-blast.phi-base.org/ PHI-base 4.10 protein sequences with the flag -evalue 1.0e-5
```{r}

# Set the file path to your .tsv file
file_path <- "/mnt/data/barry/projects/Schroeter_Pcinnamomi/PcEffector/PHI-BLAST-std_tsv_report.tsv"

# Read the .tsv file into a data frame, excluding rows containing "# "
data <- read.delim(file_path, sep = "\t", header = FALSE)
data <- data[!grepl("# ", data$V1), , drop = FALSE]

# Create a new data frame to store the transformed data
transformed_data <- data.frame(matrix(ncol = 12, nrow = nrow(data) / 12, byrow = TRUE), row.names = NULL)

# Assign the values to the respective columns in the transformed data
transformed_data$query_id <- data$V1[seq(1, nrow(data), 12)]
transformed_data$subject_id <- data$V1[seq(2, nrow(data), 12)]
transformed_data$percent_identity <- as.numeric(data$V1[seq(3, nrow(data), 12)])
transformed_data$alignment_length <- as.numeric(data$V1[seq(4, nrow(data), 12)])
transformed_data$mismatches <- as.numeric(data$V1[seq(5, nrow(data), 12)])
transformed_data$gap_opens <- as.numeric(data$V1[seq(6, nrow(data), 12)])
transformed_data$q_start <- as.numeric(data$V1[seq(7, nrow(data), 12)])
transformed_data$q_end <- as.numeric(data$V1[seq(8, nrow(data), 12)])
transformed_data$s_start <- as.numeric(data$V1[seq(9, nrow(data), 12)])
transformed_data$s_end <- as.numeric(data$V1[seq(10, nrow(data), 12)])
transformed_data$evalue <- data$V1[seq(11, nrow(data), 12)]
transformed_data$bit_score <- as.numeric(data$V1[seq(12, nrow(data), 12)])

transformed_data <- transformed_data[, -(1:12)]

phiunique <- unique(transformed_data$query_id)
```

###
bdCAN analysis
###

https://bcb.unl.edu/dbCAN2/blast.php using allpcgenes.fasta

check all boxes except substrate prediction
HMMER: dbCAN (E-Value < 1e-15, coverage > 0.35)
DIAMOND: CAZy (E-Value < 1e-102
HMMER: dbCAN-sub (E-Value < 1e-15, coverage > 0.35)

Jinfang Zheng and others, dbCAN3: automated carbohydrate-active enzyme and substrate annotation, 
Nucleic Acids Research, 2023;, gkad328, https://doi.org/10.1093/nar/gkad328

```{r}
dbcanresults <- read_tsv("dbCAN_allpcgenes.txt") 
colnames(dbcanresults)[colnames(dbcanresults) == "Gene ID"] <- "protein_id"

dbcan <- dbcanresults %>%
  filter(`#ofTools` >= 2)

#which CAZymes have signal peptides
dbcansecreted <- dplyr::right_join(detectedsecreted.id, dbcan, by = 'protein_id')

dbcansecreted2 <- dbcan %>%
  filter(protein_id %in% detectedsecreted.id$protein_id)

dbcansecreted2$Signalp <- "Y"

head(dbcansecreted)

#435 predicted CAZymes version dbCAN3.0
#paper predicted 468 they had HMMER/Hotpep/DIAMOND  Hotpep replaced with eCAMI obsolete version dbCAN2.0
#227 predicted CAZymes have signal peptides
```
```{r}
effectorliterature <- as.data.frame(c(knowneffectors$protein_id, PFAM$))
colnames(knowneffectors)[colnames(knowneffectors) == "c(nlp.df$protein_id, crn.df$protein_id, rxlr.df$protein_id)"] <- "protein_id"
```

```{r}
#save .RDS including predicted secreted, known/predicted effectors, and CAZymes to use in PcDESeq2.rmd

#save(allsecreted.id, predictedeffectors.id, predictedeffectors2.id, knowneffectors, dbcansecreted, file = "secrete.effect.RData")

```

```{r}
ora_res <- as.data.frame(enricher(gene = module_genes_list_early[[module.color]],
                                    universe = bg,
                                    maxGSSize = 5000,
                                    TERM2GENE = genesets,
                                    pAdjustMethod = "fdr",
                                    pvalueCutoff = 1,
                                    qvalueCutoff = 1))
```

## Session information

```{r,session}

sessionInfo()

```
